---
title: "Press Conference Descriptive Statistics"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
    latex_engine: xelatex
    keep_tex: false
header-includes:
  - \usepackage{booktabs}
  - \usepackage{float}
  - \usepackage{longtable}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo      = FALSE,
  warning   = FALSE,
  message   = FALSE,
  fig.align = "center",
  out.width = "100%"
)
```

```{r environment}
rm(list = ls())

library(arrow)
library(tidyverse)
library(lubridate)
library(scales)
library(patchwork)
library(knitr)
library(kableExtra)
source("../../project_paths.R")

path_da  <- paste0(media_path('data', '02-conferences', 'raw', 'attendance'), '/')
fig_path <- paste0(path_da, 'figures/')
dir.create(fig_path, showWarnings = FALSE, recursive = TRUE)
```

```{r load-data}
panel <- read_parquet(paste0(path_da, 'panel_attendance.parquet.gzip'))

panel <- panel |>
  mutate(fecha = as.Date(fecha))

lottery_date <- as.Date("2022-04-12")
```

```{r helpers}
my_theme <- function() {
  theme_minimal() +
    theme(
      plot.title    = element_text(hjust = 0.5, face = "bold"),
      plot.subtitle = element_text(hjust = 0.5),
      axis.text.x   = element_text(angle = 45, hjust = 1)
    )
}

panel_sep_theme <- function() {
  theme(
    panel.border = element_rect(colour = "grey70", fill = NA, linewidth = 0.6),
    plot.background = element_rect(fill = "white", colour = NA),
    plot.margin = margin(10, 12, 10, 12)
  )
}

save_plot <- function(p, filename, w = 12, h = 7) {
  ggsave(paste0(fig_path, filename), plot = p, width = w, height = h, dpi = 150)
}
```

# Brief explanation

Given the fact that I had already, for the first delivery, presented the empirical strategy and the data I was going to use, in this delivery I will show some descriptive statistics to show the work I have been doing over the weeks and further evaluate the feasibility of the project.

## Conference Attendance Over Time

After some painful he three panels below track the evolution of total attendance over the full sample period. The vertical dashed line marks the lottery cutoff date (2022-04-12).

**Panel A** — Daily total attendance plots the raw day-to-day variation in the number of people present at each mañanera. The LOESS smoother (span = 0.15) highlights medium-run trends without imposing a specific functional form.

**Panel B** — Monthly average attendance aggregates the data at the monthly level, reducing short-term volatility and making broader shifts around the cutoff easier to observe.

A few patterns emerge from the figure. First, attendance exhibits several clear structural shifts over time. These are largely driven by institutional factors—such as changes in the venue’s overall capacity or the relocation of conferences to other states—rather than by changes in demand to attend the press conferences themselves. The large increase toward the end of the series corresponds to the beginning of President Claudia Sheinbaum’s administration, when capacity expanded while the lottery system remained in place during the first year.

More importantly for the design, the daily series suggests that there is no meaningful change in venue capacity around the date when the lottery rule changed. In other words, the institutional change in the allocation mechanism does not appear to coincide with a change in the number of available seats. This implies that the shift in the assignment mechanism primarily affects who is able to ask a question, rather than how many people can attend.

```{r fig1, fig.height=16, fig.width=12}
conf_daily <- panel |>
  distinct(fecha, n_total_attend, n_reporteros, n_youtubers) |>
  arrange(fecha)

# 1a. Daily total attendance
p1a <- ggplot(conf_daily, aes(fecha, n_total_attend)) +
  geom_line(colour = "steelblue", linewidth = 0.4, alpha = 0.7) +
  geom_smooth(method = "loess", span = 0.15, colour = "firebrick",
              se = FALSE, linewidth = 0.9) +
  geom_vline(xintercept = lottery_date, linetype = "dashed",
             colour = "firebrick", linewidth = 0.7) +
  annotate("text", x = lottery_date + 30,
           y = max(conf_daily$n_total_attend, na.rm = TRUE) * 0.95,
           label = "Lottery cutoff\n(2022-04-12)", colour = "firebrick",
           hjust = 0, size = 3) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  scale_y_continuous(labels = comma) +
  labs(title = "Daily Total Attendance at Mañaneras",
       x = NULL, y = "Total attendees") +
  my_theme()

# 1b. Monthly average attendance
conf_monthly <- conf_daily |>
  mutate(month_date = floor_date(fecha, "month")) |>
  group_by(month_date) |>
  summarise(avg_attend = mean(n_total_attend, na.rm = TRUE), .groups = "drop")

p1b <- ggplot(conf_monthly, aes(month_date, avg_attend)) +
  geom_line(colour = "steelblue", linewidth = 0.8) +
  geom_point(colour = "steelblue", size = 1.5) +
  geom_vline(xintercept = lottery_date, linetype = "dashed",
             colour = "firebrick", linewidth = 0.7) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  scale_y_continuous(labels = comma) +
  labs(title = "Monthly Average Attendance",
       x = NULL, y = "Avg. attendees") +
  my_theme()

# 1c. Reporters vs YouTubers (monthly)
conf_monthly_types <- conf_daily |>
  mutate(month_date = floor_date(fecha, "month")) |>
  group_by(month_date) |>
  summarise(
    Reporters  = mean(n_reporteros, na.rm = TRUE),
    YouTubers  = mean(n_youtubers,  na.rm = TRUE),
    .groups = "drop"
  ) |>
  pivot_longer(c(Reporters, YouTubers), names_to = "type", values_to = "avg_count")

p1c <- ggplot(conf_monthly_types, aes(month_date, avg_count, colour = type)) +
  geom_line(linewidth = 0.8) +
  geom_vline(xintercept = lottery_date, linetype = "dashed",
             colour = "firebrick", linewidth = 0.7) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  scale_y_continuous(labels = comma) +
  scale_colour_manual(values = c("Reporters" = "steelblue", "YouTubers" = "darkorange")) +
  labs(title = "Monthly Avg. Reporters vs YouTubers",
       x = NULL, y = "Avg. count", colour = NULL) +
  my_theme()

fig1 <- (p1a / p1b) +
  plot_annotation(
    title = "Conference Attendance Over Time",
    theme = theme(plot.title = element_text(hjust = 0.5, face = "bold"))
  )

save_plot(fig1, "01_conference_attendance_timeseries.png", w = 12, h = 16)
fig1
```


\newpage

# Outlet-Level Attendance Rates

These plots shift the unit of analysis from the conference to the media outlet. For each outlet, the attendance rate is defined as the share of conferences it attended, computed separately for the pre- and post-lottery periods.

**Panel A** — Top 30 outlets identifies the outlets with the highest overall attendance rates. A clear pattern emerges: large outlets, regardless of ideology or political alignment, are those that attend the conference most frequently. Reforma (oppositional, right-leaning), El Universal (oppositional), La Jornada (aligned, left-leaning), and Grupo Cantón (aligned) are all among the outlets with the highest attendance. Part of this pattern likely reflects resource differences. Attending the conference is costly—it requires assigning reporters, paying salaries, and dedicating time that could otherwise be spent covering other events—so larger organizations are better positioned to participate consistently.

Given this structure, the presence of a lottery and the ability to control for attendance make this setting particularly useful empirically. Conditional on being present at the conference, the lottery introduces within-day exogenous variation in which outlet is selected to ask a question, which I will leverage in the analysis.

**Panel B** — Pre vs. post scatter plots each outlet according to its attendance rate before and after the lottery. Points on the 45° dashed line indicate no change in attendance behavior; points above the line correspond to outlets that attended more frequently after the lottery, while points below indicate outlets that attended less often. This visualization offers a simple first-pass assessment of whether the lottery altered the selection of outlets present at the conference. The analysis can be further refined once additional metadata on outlets becomes available—for example, information on ideology, political alignment, or outlet size.

**Panel C** — Density plot shows the full distribution of outlet attendance rates in each period. There is a small but noticeable rightward shift in the post-lottery distribution that could suggest that outlets that previously attended infrequently were able to increase their presence or were more incentivized to do so since the probability of asking a question increased. 

```{r outlets-prep}
n_conferences_total <- n_distinct(panel$fecha)
n_conf_bef          <- panel |> filter(fecha < lottery_date)  |> distinct(fecha) |> nrow()
n_conf_aft          <- panel |> filter(fecha >= lottery_date) |> distinct(fecha) |> nrow()

outlet_rates <- panel |>
  group_by(outlet_final) |>
  summarise(
    total_att  = sum(attended, na.rm = TRUE),
    att_bef    = sum(attended[fecha < lottery_date],  na.rm = TRUE),
    att_aft    = sum(attended[fecha >= lottery_date], na.rm = TRUE),
    .groups    = "drop"
  ) |>
  mutate(
    rate_total = total_att / n_conferences_total,
    rate_bef   = att_bef  / n_conf_bef,
    rate_aft   = att_aft  / n_conf_aft
  )
```

```{r fig2a, fig.height=9, fig.width=11}
top30 <- outlet_rates |>
  slice_max(total_att, n = 30) |>
  mutate(outlet_final = fct_reorder(outlet_final, total_att))

p2a <- ggplot(top30, aes(total_att, outlet_final)) +
  geom_col(fill = "steelblue") +
  scale_x_continuous(labels = comma) +
  labs(title = "Top 30 Outlets by Total Conferences Attended",
       x = "Total conferences attended", y = NULL) +
  my_theme() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

save_plot(p2a, "02a_top30_outlets.png", w = 11, h = 9)
p2a
```

```{r fig2b, fig.height=8, fig.width=9}
p2b <- ggplot(outlet_rates, aes(rate_bef, rate_aft)) +
  geom_point(colour = "steelblue", alpha = 0.5, size = 1.8) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey40") +
  scale_x_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1)) +
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1)) +
  labs(title = "Outlet Attendance Rate: Pre vs Post Lottery",
       subtitle = "Dashed line = no change; points above = more attendance after lottery",
       x = "Attendance rate before 2022-04-12",
       y = "Attendance rate from 2022-04-12") +
  my_theme()

save_plot(p2b, "02b_lottery_scatter.png", w = 9, h = 8)
p2b
```

```{r fig2c, fig.height=6, fig.width=10}
rates_long <- outlet_rates |>
  select(outlet_final, rate_bef, rate_aft) |>
  pivot_longer(c(rate_bef, rate_aft),
               names_to  = "period",
               values_to = "rate") |>
  mutate(period = recode(period,
                         rate_bef = "Pre-lottery (before 2022-04-12)",
                         rate_aft = "Post-lottery (from 2022-04-12)"))

p2c <- ggplot(rates_long, aes(rate, fill = period)) +
  geom_density(alpha = 0.4, colour = NA) +
  scale_x_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1)) +
  scale_fill_manual(values = c("Pre-lottery (before 2022-04-12)"  = "steelblue",
                               "Post-lottery (from 2022-04-12)" = "darkorange")) +
  labs(title = "Distribution of Outlet Attendance Rates",
       x = "Attendance rate", y = "Density", fill = NULL) +
  my_theme()

save_plot(p2c, "02c_rate_distribution.png", w = 10, h = 6)
p2c
```

\newpage

# Summary Statistics

The table below reports summary statistics for the main variables in the panel. It includes the number of observations and missing values, as well as standard measures of central tendency and dispersion.

Variables with (Conference) correspond to conference-level counts, while (Outlet) are outlet-level measures of participation. The remaining statistics (mean, standard deviation, and the five-number summary) provide a compact description of the distribution of each variable.

```{r summary-table}
panel <- panel |> mutate(n_other = ifelse(n_total_attend - n_journalists 
                         -n_reporteros - n_youtubers < 0, 0, 
                         n_total_attend - n_journalists - 
                           n_reporteros -n_youtubers)) 

stat_cols <- c("n_journalists", "attended", "n_total_attend",
               "n_reporteros", "n_youtubers",
               "total_attendance", "attendance_bef", "attendance_after")

stats_conf <- c("n_total_attend", 'n_other', 
               "n_reporteros", "n_youtubers")

summary_conf <- panel |> select(fecha, all_of(stats_conf)) |> 
  distinct(fecha, .keep_all = T) |> select(-fecha)|>
  pivot_longer(everything(), names_to = "variable") |>
  group_by(variable) |>
  summarise(
    n      = sum(!is.na(value)),
    n_miss = sum(is.na(value)),
    mean   = mean(value,           na.rm = TRUE),
    sd     = sd(value,             na.rm = TRUE),
    min    = min(value,            na.rm = TRUE),
    p25    = quantile(value, 0.25, na.rm = TRUE),
    median = median(value,         na.rm = TRUE),
    p75    = quantile(value, 0.75, na.rm = TRUE),
    max    = max(value,            na.rm = TRUE),
    .groups = "drop"
  )

stats_outlet <- c("total_attendance", "attendance_bef", 
                  "attendance_after")

summary_out <- panel |> select(outlet_final, all_of(stats_outlet)) |> 
  distinct(outlet_final, .keep_all = T) |> select(-outlet_final)|>
  pivot_longer(everything(), names_to = "variable") |>
  group_by(variable) |>
  summarise(
    n      = sum(!is.na(value)),
    n_miss = sum(is.na(value)),
    mean   = mean(value,           na.rm = TRUE),
    sd     = sd(value,             na.rm = TRUE),
    min    = min(value,            na.rm = TRUE),
    p25    = quantile(value, 0.25, na.rm = TRUE),
    median = median(value,         na.rm = TRUE),
    p75    = quantile(value, 0.75, na.rm = TRUE),
    max    = max(value,            na.rm = TRUE),
    .groups = "drop"
  )

summary_tbl <- rbind(summary_conf, summary_out)
  

write_csv(summary_tbl, paste0(path_da, 'descriptive_stats_summary.csv'))

summary_tbl |> 
  mutate(variable = case_when(variable == "attendance_after" ~ "Attendance After Lottery (Outlet)",
                              variable == "attendance_bef" ~ "Att. Before Lottery (Outlet)",
                              variable == "total_attendance" ~ "Total Attendance (Outlet)",
                              variable == "n_other" ~ "Total Other Types (Conference)",
                              variable == "n_reporteros" ~ "Total Reporters (Conference)",
                              variable == "n_youtubers" ~ "Total Youtubers (Conference)", 
                              variable == "n_total_attend" ~ "Total Attendance (Conference)")) |>
  rename(
    Variable = variable,
    N        = n,
    Missing  = n_miss,
    Mean     = mean,
    SD       = sd,
    Min      = min,
    P25      = p25,
    Median   = median,
    P75      = p75,
    Max      = max
  ) |>
  mutate(across(where(is.double), \(x) round(x, 2))) |>
  kbl(
    format    = "latex",
    booktabs  = TRUE,
    longtable = FALSE,
    caption   = "Descriptive Statistics for Panel Variables",
    label     = "tab:desc-stats",
    align     = c("l", rep("r", 9))
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 9
  ) |>
  column_spec(1, bold = TRUE) |>
  add_header_above(c(" " = 3, "Distribution" = 7))
```


\newpage

# Government Advertising Contracts (COMSOC)

This section describes the structure and evolution of government advertising
expenditure sourced from CompraNet/COMSOC records. The data cover contracts
between the federal government and media outlets. A few things were important to understand:

The first thing to notice is that AMLO decreased by almost 300\% the total spending in comparison with the last year of Peña Nieto's tenure.



```{r comsoc-setup}
library(readxl)

path_comsoc <- paste0(media_path('data', '06-outcomes', 'comsoc'), '/')
fig_path    <- paste0(path_comsoc, 'figures/')
dir.create(fig_path, showWarnings = FALSE, recursive = TRUE)

fix_dates <- function(df) {
  df |>
    mutate(
      fecha.de.contratopedido = if_else(
        !is.na(year.contrato) & !is.na(month.contrato) & !is.na(day.contrato),
        make_date(year.contrato, month.contrato, day.contrato),
        as.Date(fecha.de.contratopedido)
      ),
      fecha.de.gasto = if_else(
        !is.na(year.gasto) & !is.na(month.gasto) & !is.na(day.gasto),
        make_date(year.gasto, month.gasto, day.gasto),
        as.Date(fecha.de.gasto)
      )
    )
}

df_c      <- read_parquet(paste0(path_comsoc, 'contracts_2018_2024.parquet')) |> fix_dates()
df_c_full <- read_parquet(paste0(path_comsoc, 'contracts_2012_2024.parquet')) |> fix_dates()

amlo_date <- as.Date("2019-01-01")

election_events <- tibble(
  date  = as.Date(c("2021-06-06", "2022-04-10", "2024-06-02")),
  label = c("Intermediate elections", "Presidential Recall Referendum", "General Election"),
  vjust = c(1.5, 5.5, 1.5)
)
```


```{r total-spending}


df_c_agg_yearly <- df_c_full |>
  filter(between(year.gasto, 2012, 2025)) |>
  group_by(year.gasto) |>
  summarise(costo = sum(costo, na.rm = TRUE)) |>
  ungroup()

ggplot(df_c_agg_yearly, aes(x = year.gasto, y = costo)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_y_continuous(labels = label_comma()) +
  labs(
    title = "Yearly Spending",
    x = NULL,
    y = "Total Spent (Mexican Pesos)"
  ) +
  theme_minimal(base_size = 13)

```


```{r comsoc-helpers}
add_events <- function(p, events) {
  for (i in seq_len(nrow(events))) {
    p <- p +
      geom_vline(xintercept = events$date[i], linetype = "dotted",
                 colour = "grey40", linewidth = 0.6) +
      annotate("text", x = events$date[i] + 25, y = Inf,
               vjust = events$vjust[i], hjust = 0,
               label = events$label[i], colour = "grey40", size = 2.5)
  }
  p
}

ts_count_weekly <- function(data, date_col, date_label, year_min, year_max,
                            vline = NULL) {
  p <- data |>
    filter(!is.na(.data[[date_col]])) |>
    mutate(.week = floor_date(.data[[date_col]], "week", week_start = 1)) |>
    count(.week) |>
    filter(.week >= as.Date(paste0(year_min, "-01-01")),
           .week <= as.Date(paste0(year_max, "-12-31"))) |>
    ggplot(aes(.week, n)) +
    geom_col(fill = "#1F78B4", width = 5) +
    scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
    scale_y_continuous(labels = comma) +
    labs(title = paste("Weekly contracts —", date_label),
         x = NULL, y = "# contracts") +
    my_theme()

  p <- add_events(p, election_events)

  if (!is.null(vline))
    p <- p +
      geom_vline(xintercept = vline, linetype = "dashed",
                 colour = "grey20", linewidth = 0.7) +
      annotate("text", x = vline + 25, y = Inf, vjust = 9.5, hjust = 0,
               label = "AMLO takes office", colour = "grey20", size = 2.5)
  p
}

ts_costo_panels <- function(data, date_col, date_label, year_min, year_max,
                            vline = NULL) {
  d <- data |>
    filter(!is.na(.data[[date_col]]), !is.na(costo)) |>
    mutate(.date = .data[[date_col]])

  quarterly <- d |>
    mutate(.q     = quarter(.date),
           .year  = year(.date),
           .qdate = yq(paste0(.year, ": Q", .q))) |>
    filter(.year >= year_min, .year <= year_max) |>
    group_by(.qdate, .year) |>
    summarise(total = sum(costo, na.rm = TRUE), .groups = "drop")

  monthly <- d |>
    mutate(.month = floor_date(.date, "month"),
           .year  = year(.date)) |>
    filter(.year >= year_min, .year <= year_max) |>
    group_by(.month, .year) |>
    summarise(total = sum(costo, na.rm = TRUE), .groups = "drop")

  add_vline <- function(p) {
    p <- add_events(p, election_events)
    if (is.null(vline)) return(p)
    p +
      geom_vline(xintercept = vline, linetype = "dashed",
                 colour = "grey20", linewidth = 0.7) +
      annotate("text", x = vline + 25, y = Inf, vjust = 9.5, hjust = 0,
               label = "AMLO takes office", colour = "grey20", size = 2.5)
  }

  p_quarterly <- add_vline(
    ggplot(quarterly, aes(.qdate, total)) +
      geom_col(fill = "#1F78B4") +
      geom_line(aes(group = 1), colour = "grey30", linewidth = 0.5) +
      scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
      scale_y_continuous(labels = dollar_format(prefix = "$")) +
      labs(title = paste("Quarterly spending —", date_label),
           x = NULL, y = "Total costo (MXN)") +
      my_theme()
  )

  p_quarterly_log <- add_vline(
    ggplot(quarterly, aes(.qdate, total)) +
      geom_col(fill = "#1F78B4") +
      geom_line(aes(group = 1), colour = "grey30", linewidth = 0.5) +
      scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
      scale_y_log10(labels = dollar_format(prefix = "$")) +
      labs(title = paste("Quarterly spending (log) —", date_label),
           x = NULL, y = "Total costo (MXN, log scale)") +
      my_theme()
  )

  p_monthly <- add_vline(
    ggplot(monthly, aes(.month, total)) +
      geom_col(fill = "#1F78B4") +
      scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
      scale_y_continuous(labels = dollar_format(prefix = "$")) +
      labs(title = paste("Monthly spending —", date_label),
           x = NULL, y = "Total costo (MXN)") +
      my_theme()
  )

  p_monthly_log <- add_vline(
    ggplot(monthly, aes(.month, total)) +
      geom_col(fill = "#1F78B4") +
      scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
      scale_y_log10(labels = dollar_format(prefix = "$")) +
      labs(title = paste("Monthly spending (log) —", date_label),
           x = NULL, y = "Total costo (MXN, log scale)") +
      my_theme()
  )

  list(quarterly     = p_quarterly,
       quarterly_log = p_quarterly_log,
       monthly       = p_monthly,
       monthly_log   = p_monthly_log)
}
```

## Weekly Contract Volume

The weekly bar charts below report the number of contracts signed or expensed in each week. Using both date fields—contract date and expense date—is important because administrative lags can generate substantial differences between the date a contract is formally recorded and the date payment is actually disbursed. The dotted vertical lines mark the three major electoral events in the period, while the dashed line marks the beginning of the AMLO administration.

The 2012–2024 panel provides a longer baseline, which helps distinguish spending patterns during the AMLO administration from those observed under previous governments.

Several patterns stand out. First, contract assignment does not appear to follow a mechanical seasonal pattern, such as concentration at the beginning of the year or at regular quarterly intervals. Instead, spending spikes occur at different points within each year, and these peaks do not align systematically across years. Second, there is a visible increase in the number of contracts during the 2024 electoral year.

Expense date do follow a clear pattern: contracts are being paid on the second half of the year, mostly on December.

This is useful because it suggests substantial temporal variation in contract allocation, which is consistent with the possibility that spending responds to a dynamic relationship between government and media outlets shaped by repeated interactions over time. At the same time, a more disaggregated analysis is needed to determine whether this variation is concentrated among the outlets that attend the conference and also receive COMSOC contracts.

Finally, an important caveat is that this spending is not allocated exclusively by the executive office. Multiple federal agencies—such as the Ministries of Health and Finance—also use COMSOC resources for promotional purposes. As a result, part of the observed variation may reflect agency-specific spending decisions that are independent of media dissent or the executive’s strategic response to it.

```{r comsoc-fig1, fig.height=10, fig.width=12}
p_count_contrato <- ts_count_weekly(df_c, "fecha.de.contratopedido", "Contract date", 2018, 2024)
p_count_gasto    <- ts_count_weekly(df_c, "fecha.de.gasto",          "Expense date",  2018, 2024)

fig_c1 <- (p_count_contrato / p_count_gasto) +
  plot_annotation(title = "Weekly Number of Contracts (2018\u20132024)",
                  theme = theme(plot.title = element_text(hjust = 0.5, face = "bold")))

save_plot(fig_c1, "01_count_weekly.png", w = 12, h = 10)
fig_c1
```

```{r comsoc-fig1-full, fig.height=10, fig.width=12}
p_count_contrato_full <- ts_count_weekly(df_c_full, "fecha.de.contratopedido",
                                         "Contract date", 2012, 2024, vline = amlo_date)
p_count_gasto_full    <- ts_count_weekly(df_c_full, "fecha.de.gasto",
                                         "Expense date",  2012, 2024, vline = amlo_date)

fig_c1_full <- (p_count_contrato_full / p_count_gasto_full) +
  plot_annotation(title = "Weekly Number of Contracts (2012\u20132024)",
                  theme = theme(plot.title = element_text(hjust = 0.5, face = "bold")))

save_plot(fig_c1_full, "01_count_weekly_full.png", w = 12, h = 10)
fig_c1_full
```

\newpage

## Total Spending Over Time

Spending is aggregated at both the quarterly and monthly level. The two
frequency choices serve different purposes: quarterly smooths over month-end
billing cycles, while monthly preserves intra-year seasonality. Each is shown
on a linear scale.

Conclusions are similar: there seems to be enough variation both on contracts assigned and on money spent across the year.

```{r comsoc-fig2-contrato, fig.height=12, fig.width=12}
panels_costo_contrato <- ts_costo_panels(df_c, "fecha.de.contratopedido",
                                         "Contract date", 2018, 2024)

fig_c2a_total <- (panels_costo_contrato$quarterly / panels_costo_contrato$monthly) +
  plot_annotation(title = "Total Spending (2018\u20132024) \u2014 Contract date (linear scale)",
                  theme = theme(plot.title = element_text(hjust = 0.5, face = "bold")))

fig_c2a_log <- (panels_costo_contrato$quarterly_log / panels_costo_contrato$monthly_log) +
  plot_annotation(title = "Total Spending (2018\u20132024) \u2014 Contract date (log scale)",
                  theme = theme(plot.title = element_text(hjust = 0.5, face = "bold")))

save_plot(fig_c2a_total, "02a_costo_contract_date.png",     w = 12, h = 12)
save_plot(fig_c2a_log,   "02a_costo_contract_date_log.png", w = 12, h = 12)
print(fig_c2a_total)
#fig_c2a_log
```

```{r comsoc-fig2-gasto, fig.height=12, fig.width=12}
panels_costo_gasto <- ts_costo_panels(df_c, "fecha.de.gasto",
                                      "Expense date", 2018, 2024)

fig_c2b_total <- (panels_costo_gasto$quarterly / panels_costo_gasto$monthly) +
  plot_annotation(title = "Total Spending (2018\u20132024) \u2014 Expense date (linear scale)",
                  theme = theme(plot.title = element_text(hjust = 0.5, face = "bold")))

fig_c2b_log <- (panels_costo_gasto$quarterly_log / panels_costo_gasto$monthly_log) +
  plot_annotation(title = "Total Spending (2018\u20132024) \u2014 Expense date (log scale)",
                  theme = theme(plot.title = element_text(hjust = 0.5, face = "bold")))

save_plot(fig_c2b_total, "02b_costo_expense_date.png",     w = 12, h = 12)
save_plot(fig_c2b_log,   "02b_costo_expense_date_log.png", w = 12, h = 12)
print(fig_c2b_total)
#fig_c2b_log
```

```{r comsoc-fig2-full, fig.height=12, fig.width=12}
panels_costo_contrato_full <- ts_costo_panels(df_c_full, "fecha.de.contratopedido",
                                              "Contract date", 2012, 2024, vline = amlo_date)
panels_costo_gasto_full    <- ts_costo_panels(df_c_full, "fecha.de.gasto",
                                              "Expense date",  2012, 2024, vline = amlo_date)

fig_c2_full_total <- (panels_costo_contrato_full$quarterly / panels_costo_contrato_full$monthly) +
  plot_annotation(title = "Total Spending (2012\u20132024) \u2014 Contract date (linear scale)",
                  theme = theme(plot.title = element_text(hjust = 0.5, face = "bold")))

fig_c2_full_log <- (panels_costo_contrato_full$quarterly_log /
                    panels_costo_contrato_full$monthly_log) +
  plot_annotation(title = "Total Spending (2012\u20132024) \u2014 Contract date (log scale)",
                  theme = theme(plot.title = element_text(hjust = 0.5, face = "bold")))

save_plot(fig_c2_full_total, "02a_costo_contract_date_full.png",     w = 12, h = 12)
save_plot(fig_c2_full_log,   "02a_costo_contract_date_full_log.png", w = 12, h = 12)
print(fig_c2_full_total)
#fig_c2_full_log
```

\newpage

## Beneficiaries, Contract Sizes, and Product Types

```{r comsoc-clean-load}
df_c_clean <- read_parquet(paste0(path_comsoc, 'contracts_2018_2024_clean.parquet'))
```

**Top 20 beneficiaries** identifies the outlets that captured the largest share
of total advertising spend. Concentration here — a small number of outlets
receiving a disproportionate share — would be the expected pattern if government
advertising is used to reward loyal media.

```{r comsoc-fig3a, fig.height=8, fig.width=11}
top20 <- df_c_clean |>
  filter(!is.na(bene_final), !is.na(costo)) |>
  group_by(bene_final) |>
  summarise(total = sum(costo, na.rm = TRUE), .groups = "drop") |>
  slice_max(total, n = 20) |>
  mutate(bene_final = fct_reorder(bene_final, total))

p3a <- ggplot(top20, aes(total, bene_final)) +
  geom_col(fill = "#1F78B4") +
  scale_x_continuous(labels = dollar_format(prefix = "$")) +
  labs(title = "Top 20 Beneficiaries by Total Spending",
       x = "Total costo (MXN)", y = NULL) +
  my_theme() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

save_plot(p3a, "03a_top20_beneficiaries.png", w = 11, h = 8)
p3a
```

**Contract size distribution** is plotted on a log$_{10}$ scale because
government contract values span several orders of magnitude.

```{r comsoc-fig3b, fig.height=6, fig.width=10}
df_c_pos <- df_c_clean |> filter(costo > 0, !is.na(costo))

p3b <- ggplot(df_c_pos, aes(costo)) +
  geom_histogram(bins = 80, fill = "#1F78B4", colour = "white", linewidth = 0.2) +
  scale_x_log10(labels = dollar_format(prefix = "$")) +
  scale_y_continuous(labels = comma) +
  labs(title = "Distribution of Contract Sizes (log\u2081\u2080 scale)",
       x = "Contract amount \u2014 costo (MXN, log scale)", y = "Count") +
  my_theme()

save_plot(p3b, "03b_contract_size_distribution.png", w = 10, h = 6)
p3b
```


## Contracts Over Time — Top 200 Outlets by Total Spend

Restricting the sample to the 200 outlets that received the most total
advertising money focuses the time-series analysis on the outlets that are
substantively important for the study and removes the large mass of
occasional or trivial recipients that would otherwise dilute the signal.

Results hold as well, there is still enough variation and no clear patterns.

The four panels below characterise how activity among these high-value
outlets evolved:

- **Weekly contract count** shows the rhythm of contracting — whether
  the government distributed contracts in bursts (e.g., around elections)
  or at a steady rate.
- **Weekly total amount** tracks the peso value of those contracts. A
  spike in count without a proportional spike in value (or vice versa)
  indicates changes in the average contract size, which may reflect
  different procurement strategies.
- **Total spend per outlet** (bar chart, top 50 shown for legibility)
  reveals concentration within the top-200 group — whether the group is
  itself dominated by a handful of outlets or whether spending is
  relatively diffuse.
- **Contract-size distribution** for the top-200 outlets compared to the
  rest of the data shows whether high-value recipients tend to receive
  larger individual contracts or simply more frequent ones.

```{r comsoc-top200-prep}
top200_benes <- df_c_clean |>
  filter(!is.na(bene_final), !is.na(costo)) |>
  group_by(bene_final) |>
  summarise(total_costo = sum(costo, na.rm = TRUE), .groups = "drop") |>
  slice_max(total_costo, n = 200, with_ties = FALSE)

df_top200 <- df_c_clean |>
  filter(bene_final %in% top200_benes$bene_final,
         !is.na(fecha.de.contratopedido))
```

```{r comsoc-top200-weekly, fig.height=10, fig.width=12}
top200_weekly <- df_top200 |>
  mutate(.week = floor_date(fecha.de.contratopedido, "week", week_start = 1)) |>
  group_by(.week) |>
  summarise(
    n_contracts  = n(),
    total_amount = sum(costo, na.rm = TRUE),
    .groups      = "drop"
  ) |>
  filter(.week >= as.Date("2018-01-01"),
         .week <= as.Date("2024-12-31"))

p_t200_count <- ggplot(top200_weekly, aes(.week, n_contracts)) +
  geom_col(fill = "#1F78B4", width = 5) +
  geom_vline(data = election_events, aes(xintercept = date),
             linetype = "dotted", colour = "grey40", linewidth = 0.6) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  scale_y_continuous(labels = comma) +
  labs(title = "Weekly number of contracts — top 200 outlets",
       x = NULL, y = "# contracts") +
  my_theme()

for (i in seq_len(nrow(election_events))) {
  p_t200_count <- p_t200_count +
    annotate("text", x = election_events$date[i] + 25, y = Inf,
             vjust = election_events$vjust[i], hjust = 0,
             label = election_events$label[i], colour = "grey40", size = 2.5)
}

p_t200_amount <- ggplot(top200_weekly, aes(.week, total_amount)) +
  geom_col(fill = "#E31A1C", width = 5) +
  geom_vline(data = election_events, aes(xintercept = date),
             linetype = "dotted", colour = "grey40", linewidth = 0.6) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  scale_y_continuous(labels = dollar_format(prefix = "$")) +
  labs(title = "Weekly total amount — top 200 outlets",
       x = NULL, y = "Total costo (MXN)") +
  my_theme()

for (i in seq_len(nrow(election_events))) {
  p_t200_amount <- p_t200_amount +
    annotate("text", x = election_events$date[i] + 25, y = Inf,
             vjust = election_events$vjust[i], hjust = 0,
             label = election_events$label[i], colour = "grey40", size = 2.5)
}

fig_t200_ts <- (p_t200_count / p_t200_amount) +
  plot_annotation(
    title    = "Contract Activity Among Top 200 Outlets by Total Spend (2018\u20132024)",
    subtitle = "Weekly aggregates; dotted lines mark electoral events",
    theme    = theme(plot.title    = element_text(hjust = 0.5, face = "bold"),
                     plot.subtitle = element_text(hjust = 0.5))
  )

save_plot(fig_t200_ts, "04a_top200_timeseries.png", w = 12, h = 10)
fig_t200_ts
```

```{r comsoc-top200-bar, fig.height=9, fig.width=11}
top50_within200 <- top200_benes |>
  slice_max(total_costo, n = 50, with_ties = FALSE) |>
  mutate(bene_final = fct_reorder(bene_final, total_costo))

p_t200_bar <- ggplot(top50_within200, aes(total_costo, bene_final)) +
  geom_col(fill = "#1F78B4") +
  scale_x_continuous(labels = dollar_format(prefix = "$")) +
  labs(title = "Total Spend per Outlet — Top 50 Within the Top-200 Group",
       x = "Total costo (MXN, 2018\u20132024)", y = NULL) +
  my_theme() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

save_plot(p_t200_bar, "04b_top200_bar.png", w = 11, h = 9)
p_t200_bar
```

```{r comsoc-top200-dist, fig.height=6, fig.width=10}
df_c_pos_flag <- df_c_clean |>
  filter(!is.na(costo), costo > 0) |>
  mutate(group = if_else(bene_final %in% top200_benes$bene_final,
                         "Top 200 outlets", "All other outlets"))

p_t200_dist <- ggplot(df_c_pos_flag, aes(costo, fill = group)) +
  geom_density(alpha = 0.45, colour = NA) +
  scale_x_log10(labels = dollar_format(prefix = "$")) +
  scale_fill_manual(values = c("Top 200 outlets"  = "#E31A1C",
                               "All other outlets" = "#1F78B4")) +
  labs(title = "Contract-Size Distribution: Top 200 vs Rest (log\u2081\u2080 scale)",
       x = "Contract amount — costo (MXN)", y = "Density", fill = NULL) +
  my_theme()

save_plot(p_t200_dist, "04c_top200_size_dist.png", w = 10, h = 6)
p_t200_dist
```

\newpage

# YouTube Channel Outcomes

This section characterises data on the Youtube channels of 178 Mexican media outlets (radio, TV, newspapers) that attended the press conference, were assigned a contract, or both. Over these two weeks I manually deduplicated and obtained the YouTube handles of each outlet and used the YouTube API to download all the relevant videos.

Youtube views, topics of the videos, likes and comments could be used as an outcome that might inform about effects of participating in the conference on engagement, and therefore on private advertising revenue/platform revenue. 

```{r yt-setup}
PANEL_PATH <- media_path('data', '06-outcomes', 'youtube', 'panel_daily.parquet')
DIM_PATH   <- media_path('data', '06-outcomes', 'youtube', 'mexican_newspapers_youtube.xlsx')
OUT_FOLDER <- dirname(media_output_path('data', '06-outcomes', 'youtube', 'graphs', '.keep'))

dir.create(OUT_FOLDER, recursive = TRUE, showWarnings = FALSE)

panel_yt <- read_parquet(PANEL_PATH) |>
  mutate(date = as.Date(date))

dim_ch <- read_excel(DIM_PATH) |>
  select(channel_id        = channelId,
         name_page         = `name.page`,
         yt_handle,
         yt_subscriberCount,
         yt_videoCount,
         yt_viewCount,
         yt_createdAt) |>
  filter(!is.na(channel_id)) |>
  distinct(channel_id, .keep_all = TRUE) |>
  mutate(created_year = year(as.Date(substr(yt_createdAt, 1, 10))))

save_fig <- function(p, name, w = 12, h = 7) {
  ggsave(file.path(OUT_FOLDER, name), p, width = w, height = h,
         dpi = 300, bg = "white")
}
```

## Time Series: Videos, Views, and Engagement

The three-panel time series aggregates all channels into a single weekly or monthly series. This macro view helps capture overall trends in media activity on YouTube.

- **Videos published** shows that total output by media outlets increased steadily over the sample period.
- **Total views** serves as the main measure of audience reach.
- **Likes + comments** (engagement) captures active audience interaction rather than passive consumption. Engagement often responds more directly to content choices than raw view counts.

Both weekly and monthly aggregations are reported. Weekly data highlights short-run volatility and potential news-cycle effects, while the monthly series makes medium-run trends easier to observe.

Overall, the figures show that YouTube has become an increasingly important platform for media outlets over time. Both the volume of uploaded content and the level of audience activity—measured through views, likes, and comments—have grown substantially throughout the period.

```{r yt-fig1-weekly, fig.height=9, fig.width=13}
make_timeseries_plot <- function(df_agg, freq_label) {
  p1 <- ggplot(df_agg, aes(x = period, y = n_videos)) +
    geom_line(colour = "#2C7BB6", linewidth = 0.6) +
    geom_area(fill = "#2C7BB6", alpha = 0.15) +
    scale_y_continuous(labels = comma) +
    scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
    labs(title = "Videos published", x = NULL, y = "# videos") +
    my_theme() +
    panel_sep_theme()

  p2 <- ggplot(df_agg, aes(x = period, y = viewCount)) +
    geom_line(colour = "#1A9641", linewidth = 0.6) +
    geom_area(fill = "#1A9641", alpha = 0.15) +
    scale_y_continuous(labels = label_number(scale_cut = cut_short_scale())) +
    scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
    labs(title = "Total views", x = NULL, y = "views") +
    my_theme() +
    panel_sep_theme()

  df_eng <- df_agg |> mutate(engagement = likeCount + commentCount)
  p3 <- ggplot(df_eng, aes(x = period, y = engagement)) +
    geom_line(colour = "#D7191C", linewidth = 0.6) +
    geom_area(fill = "#D7191C", alpha = 0.15) +
    scale_y_continuous(labels = label_number(scale_cut = cut_short_scale())) +
    scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
    labs(title = "Likes + comments", x = NULL, y = "count") +
    my_theme() +
    panel_sep_theme()

  (p1 / p2 / p3) +
    plot_layout(heights = c(1, 1, 1)) &
    theme(panel.spacing = unit(10, "mm")) +
    plot_annotation(
      title    = sprintf("Mexican newspaper YouTube channels \u2014 %s aggregate", freq_label),
      subtitle = "All channels combined, 2018\u20132024",
      theme    = theme(plot.title = element_text(face = "bold", size = 13))
    )
}

weekly_yt <- panel_yt |>
  mutate(period = floor_date(date, "week")) |>
  group_by(period) |>
  summarise(n_videos     = sum(n_videos,     na.rm = TRUE),
            viewCount    = sum(viewCount,    na.rm = TRUE),
            likeCount    = sum(likeCount,    na.rm = TRUE),
            commentCount = sum(commentCount, na.rm = TRUE),
            .groups = "drop")

fig_yt1_weekly <- make_timeseries_plot(weekly_yt, "weekly")
save_fig(fig_yt1_weekly, "fig1_timeseries_weekly.png", w = 13, h = 9)
fig_yt1_weekly
```

```{r yt-fig1-monthly, fig.height=9, fig.width=13}
panel_yt_active <- panel_yt |> filter(n_videos > 0)
monthly_yt <- panel_yt |>
  mutate(period = floor_date(date, "month")) |>
  group_by(period) |>
  summarise(n_videos     = sum(n_videos,     na.rm = TRUE),
            viewCount    = sum(viewCount,    na.rm = TRUE),
            likeCount    = sum(likeCount,    na.rm = TRUE),
            commentCount = sum(commentCount, na.rm = TRUE),
            .groups = "drop")

fig_yt1_monthly <- make_timeseries_plot(monthly_yt, "monthly")
save_fig(fig_yt1_monthly, "fig1_timeseries_monthly.png", w = 13, h = 9)
fig_yt1_monthly
```

\newpage

## Channel-Level Distributions

The four panels below characterise the cross-sectional distribution of channels
at the time of data collection (from the YouTube Data API metadata).

- **Subscriber count** and **total channel views** (both log-scaled) reveal
  the size distribution of channels. The figures point to a large market for
  YouTube news consumption, with substantial heterogeneity in the audience size
  reached by different outlets.
- **Total videos on channel** shows that most outlets produce content at a
  sufficiently high frequency, which makes it possible to capture the
  day-to-day or weekly variation generated by their recurrent interactions with
  the conference.
- **Channel creation year** indicates when outlets entered YouTube. Most
  channels appear to be relatively old and well established on the platform.

```{r yt-fig2, fig.height=8, fig.width=13}
dim_plot <- dim_ch |> filter(!is.na(yt_subscriberCount) | !is.na(yt_videoCount))

p_subs <- ggplot(
    dim_plot |> filter(!is.na(yt_subscriberCount), yt_subscriberCount > 0),
    aes(x = yt_subscriberCount)) +
  geom_histogram(bins = 30, fill = "#2C7BB6", colour = "white", linewidth = 0.2) +
  scale_x_log10(labels = label_number(scale_cut = cut_short_scale())) +
  labs(title = "Subscriber count", x = "subscribers (log scale)", y = "channels") +
  my_theme() +
  panel_sep_theme()

p_vids <- ggplot(
    dim_plot |> filter(!is.na(yt_videoCount), yt_videoCount > 0),
    aes(x = yt_videoCount)) +
  geom_histogram(bins = 30, fill = "#1A9641", colour = "white", linewidth = 0.2) +
  scale_x_log10(labels = comma) +
  labs(title = "Total videos on channel", x = "videos (log scale)", y = "channels") +
  my_theme() +
  panel_sep_theme()

p_views <- ggplot(
    dim_plot |> filter(!is.na(yt_viewCount), yt_viewCount > 0),
    aes(x = yt_viewCount)) +
  geom_histogram(bins = 30, fill = "#D7191C", colour = "white", linewidth = 0.2) +
  scale_x_log10(labels = label_number(scale_cut = cut_short_scale())) +
  labs(title = "Total channel views", x = "views (log scale)", y = "channels") +
  my_theme() +
  panel_sep_theme()

year_counts <- dim_ch |> filter(!is.na(created_year)) |> count(created_year)

p_year <- ggplot(year_counts, aes(x = factor(created_year), y = n)) +
  geom_col(fill = "#756BB1", colour = "white", linewidth = 0.2) +
  labs(title = "Channel creation year", x = NULL, y = "channels") +
  my_theme() +
  panel_sep_theme()

fig_yt2 <- (p_subs | p_vids) / (p_views | p_year) +
  plot_layout(guides = "collect") &
  theme(panel.spacing = unit(10, "mm")) +
  plot_annotation(
    title    = "Channel-level descriptive statistics",
    subtitle = "From YouTube Data API channel metadata",
    theme    = theme(plot.title = element_text(face = "bold", size = 13))
  )

save_fig(fig_yt2, "fig2_channel_distributions.png", w = 13, h = 8)
fig_yt2
```

## Top Channels by Output and Reach

**Top 20 by total videos** identifies the most prolific publishers.

**Top 20 by total views** (Fig. 4) identifies the channels with the greatest
cumulative audience exposure over the sample period.

Overall, large media outlets tend to appear at the top of the distribution in
terms of total videos published, which is consistent with their greater
resources and production capacity. However, there is also noticeable
heterogeneity. Smaller or independent outlets—such as El Chapucero (aligned),
El Astillero (aligned), Infobae (impartial), and El Financiero (opposition)—
also accumulate substantial viewership. This suggests that some smaller outlets
may be leveraging social media platforms as a lower-cost strategy to expand
their reach, relative to more resource-intensive formats such as television or
print media.

```{r yt-fig3, fig.height=7, fig.width=11}
ch_totals_yt <- panel_yt |>
  group_by(channel_id, `name.page`) |>
  summarise(
    total_videos   = sum(n_videos,     na.rm = TRUE),
    total_views    = sum(viewCount,    na.rm = TRUE),
    total_likes    = sum(likeCount,    na.rm = TRUE),
    total_comments = sum(commentCount, na.rm = TRUE),
    active_days    = sum(published,    na.rm = TRUE),
    .groups = "drop"
  )

top_videos <- ch_totals_yt |>
  group_by(`name.page`) |>
  summarise(total_videos = sum(total_videos, na.rm = TRUE), .groups = "drop") |>
  slice_max(total_videos, n = 20, with_ties = FALSE) |>
  mutate(`name.page` = fct_reorder(`name.page`, total_videos))

fig_yt3 <- ggplot(top_videos, aes(x = total_videos, y = `name.page`)) +
  geom_col(fill = "#2C7BB6") +
  geom_text(aes(label = comma(total_videos)), hjust = -0.1, size = 3) +
  scale_x_continuous(labels = comma, expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Top 20 channels by total videos published (2018\u20132024)",
       x = "total videos", y = NULL) +
  my_theme()

save_fig(fig_yt3, "fig3_top20_videos.png", w = 11, h = 7)
fig_yt3
```

```{r yt-fig4, fig.height=7, fig.width=11}
top_views <- ch_totals_yt |>
  group_by(`name.page`) |>
  summarise(total_views = sum(total_views, na.rm = TRUE), .groups = "drop") |>
  slice_max(total_views, n = 20, with_ties = FALSE) |>
  mutate(`name.page` = fct_reorder(`name.page`, total_views))

fig_yt4 <- ggplot(top_views, aes(x = total_views, y = `name.page`)) +
  geom_col(fill = "#1A9641") +
  geom_text(aes(label = label_number(scale_cut = cut_short_scale())(total_views)),
            hjust = -0.1, size = 3) +
  scale_x_continuous(labels = label_number(scale_cut = cut_short_scale()),
                     expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Top 20 channels by total views (2018\u20132024)",
       x = "total views", y = NULL) +
  my_theme()

save_fig(fig_yt4, "fig4_top20_views.png", w = 11, h = 7)
fig_yt4
```

\newpage

## Publishing Frequency Heatmap

The heatmap displays total videos published by all channels combined for every
day-of-week × week-of-year cell, faceted by year. It serves two descriptive
purposes:

In general, the plot shows that media outlets follow a clear weekday cycle,
similar to the rhythm of the press conference: most videos are posted during
weekdays rather than weekends. It also highlights a substantial increase in
posting activity over time. In particular, 2024 stands out as the year with the
highest level of YouTube activity, which is notable given that it also
corresponds to significant variation in advertising spending and the continued
presence of the conference.

```{r yt-fig6, fig.height=10, fig.width=13}
heat_df <- panel_yt |>
  mutate(
    year = year(date),
    week = isoweek(date),
    dow  = wday(date, label = TRUE, abbr = TRUE, week_start = 1)
  ) |>
  group_by(year, week, dow) |>
  summarise(n_videos = sum(n_videos, na.rm = TRUE), .groups = "drop")

fig_yt6 <- ggplot(heat_df, aes(x = week, y = fct_rev(dow), fill = n_videos)) +
  geom_tile(colour = "white", linewidth = 0.3) +
  facet_wrap(~year, ncol = 1) +
  scale_fill_distiller(palette = "YlOrRd", direction = 1,
                       labels = comma, name = "videos") +
  scale_x_continuous(breaks = c(1, 13, 26, 39, 52),
                     labels = c("Jan", "Apr", "Jul", "Oct", "Dec")) +
  labs(title = "Publishing frequency heatmap \u2014 all channels combined",
       subtitle = "Total videos published by day-of-week and week-of-year",
       x = NULL, y = NULL) +
  theme(strip.text       = element_text(face = "bold"),
        panel.spacing.y  = unit(4, "pt"),
        legend.position  = "bottom")

save_fig(fig_yt6, "fig6_publishing_heatmap.png", w = 13, h = 10)
fig_yt6
```

\newpage

## Summary Statistics — YouTube Panel

Two tables are reported. **Table A** covers channel-level totals (summed across
the full 2018--2024 panel per channel). These are the units of observation in
any cross-sectional analysis and determine the range of heterogeneity that
fixed effects or controls need to absorb. **Table B** covers per-video metrics
computed on active days only (days with at least one video published), which is
the relevant sample for engagement-per-video outcome specifications.

```{r yt-summary-table}
stats_vars <- c("total_videos", "total_views", "total_likes",
                "total_comments", "active_days")

yt_ch_summary <- ch_totals_yt |>
  select(all_of(stats_vars)) |>
  pivot_longer(everything(), names_to = "variable") |>
  group_by(variable) |>
  summarise(
    n      = sum(!is.na(value)),
    n_miss = sum(is.na(value)),
    mean   = mean(value,           na.rm = TRUE),
    sd     = sd(value,             na.rm = TRUE),
    min    = min(value,            na.rm = TRUE),
    p25    = quantile(value, 0.25, na.rm = TRUE),
    median = median(value,         na.rm = TRUE),
    p75    = quantile(value, 0.75, na.rm = TRUE),
    max    = max(value,            na.rm = TRUE),
    .groups = "drop"
  )

pv_vars <- c("viewCount_per_video", "likeCount_per_video", "commentCount_per_video")

yt_pv_summary <- panel_yt_active |>
  select(all_of(pv_vars)) |>
  pivot_longer(everything(), names_to = "variable") |>
  group_by(variable) |>
  summarise(
    n      = sum(!is.na(value)),
    n_miss = sum(is.na(value)),
    mean   = mean(value,           na.rm = TRUE),
    sd     = sd(value,             na.rm = TRUE),
    min    = min(value,            na.rm = TRUE),
    p25    = quantile(value, 0.25, na.rm = TRUE),
    median = median(value,         na.rm = TRUE),
    p75    = quantile(value, 0.75, na.rm = TRUE),
    max    = max(value,            na.rm = TRUE),
    .groups = "drop"
  )

fmt_tbl <- function(df, caption, label) {
  df |>
    rename(Variable = variable, N = n, Missing = n_miss,
           Mean = mean, SD = sd, Min = min, P25 = p25,
           Median = median, P75 = p75, Max = max) |>
    mutate(across(where(is.double), \(x) round(x, 1))) |>
    kbl(
      format    = "latex",
      booktabs  = TRUE,
      longtable = FALSE,
      caption   = caption,
      label     = label,
      align     = c("l", rep("r", 9))
    ) |>
    kable_styling(
      latex_options = c("hold_position", "scale_down"),
      font_size     = 9
    ) |>
    column_spec(1, bold = TRUE) |>
    add_header_above(c(" " = 3, "Distribution" = 7))
}

fmt_tbl(yt_ch_summary,
        "Table A: Channel-Level Totals (2018--2024)",
        "tab:yt-ch-stats")
```

```{r yt-summary-table-pv}
fmt_tbl(yt_pv_summary,
        "Table B: Per-Video Metrics (Active Days Only)",
        "tab:yt-pv-stats")
```

# Conclusion and Next Steps (Data)

Overall, the data suggest that there is sufficient temporal variation—at least at
the weekly level, and potentially at the daily level—in both the public
advertising outcomes (dissent) and the private advertising proxies. This level
of variation is encouraging for the empirical strategy, as it provides the type
of within-period fluctuations needed to study the dynamic relationship between
government and media outlets.

This document also reflects part of the data work carried out over the past few
weeks. A significant portion of the effort has involved manually deduplicating
media outlets across datasets. This process is not trivial, as outlets often
appear under different names or legal entities (*razones sociales*), which
requires careful verification. In addition, substantial manual work has gone
into identifying outlet metadata—such as YouTube handles and newspaper
websites—through direct searches on YouTube and Google.

There are still several steps to complete. First, I need to further improve the
deduplication of the COMSOC contracts in order to merge them cleanly with both
the YouTube dataset and the conference attendance dataset. I also plan to
collect additional outcome measures, including daily Google Trends data for
news outlets and Twitter activity.

Most importantly, I have already obtained the transcripts of the *Mañanera*
press conferences. From these scripts, I plan to extract participation
information—specifically which outlets participated and how—using an LLM,
followed by manual verification using the corresponding YouTube videos.
Fortunately, journalists typically introduce themselves and their media outlet
before asking a question, which should make automated extraction feasible. Once
this participation dataset is constructed, I will be able to begin estimating
regressions using the outcomes already assembled.

Finally, an important next step is to collect richer outlet-level metadata,
including measures of ideology and political alignment, as well as other
relevant contextual variables such as presidential approval. These additions
would allow for a more nuanced analysis of how different types of outlets
interact with the government and respond to the incentives created by the
conference.
